Malacology: A Programmable Storage System

### 2016-10-15 paper/paper.tex
:intro tweaks

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propogate cluster metadata, recover
from failure, and to make trade-offs for latency/throughput using leases.
\end{abstract}

@@

A number of recent developments have disrupted traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

These three trends put evolutionary pressure on storage systems and raise the
question of whether there are principles that storage systems designers can follow
to evolve storage systems efficiently and without jeopardizing years of
code-hardening and performance optimization efforts that are important for
users to continue to entrust their data to the storage system.

In this paper we investigate an approach that focuses on generalizing and
composing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. We will refer to this programmability as
\emph{programmable storage}, which differs from \emph{active storage} (the
injection and execution of arbitrary code in a storage system) and
\emph{software-defined storage} (the control of thin-provisioning of storage).

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the Malacology in Ceph, a popular open source software storage stack.
We choose Ceph to demonstrate the concept of programmable storage because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and object interfaces
provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, Malacology is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

{\it \textbf{The definition a programmable storage system}} to be "a storage
system that allows exisiting abstractions to be re-used and composed to form
new data and metadata management services whose operation is orothogonal to
system correctness".  Our prototype includes (but is not limited to) the
following building blocks:

@@

\item An interface for managing strongly-consistent time-varying service
metadata.

\item An interface for installing and evolving domain-specific, cluster-wide data
interfaces.

\item An interface for managing access to shared resources using a variety of
optimization strategies.

\item An interface for load balancing resources across the cluster.

\item A durability interface for persisting policies using the underyling storage
stack's object store

@@

\item A high-performance distributed shared log service called ZLog, that is an
implementation of CORFU~\cite{balakrishnan_corfu_2012}

@@

\section{Application-Specific Storage Stacks
}\label{highly-tailored-and-application-specifc-storage-systems}

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by providing a
variety of interfaces and tunable parameters. Unfortunately, managing competing
forces in these systems is difficult and users
want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

@@

To demonstrate this trend in storage systems we examine the state of
programmability in Ceph. Something of a storage swiss army knife, Ceph simultaneously supports file,
block, and object interfaces on a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
programmability; the OSDs support domain-specific object interfaces that are
implemented by composing existing low-level storage interfaces that execute
atomically. These interfaces are
written in C++ and are statically loaded into the system.

@@

The Ceph community provides empirical evidence that developers embrace
programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
growth in the production use of domain-specific interfaces in the Ceph
community since 2010. In that figure, classes are functional groupings of methods on storage
objects.  What is most remarkable is that this trend contradicts the notion
that API changes are a burden for users.  Rather it appears that a gap in
existing interfaces are being addressed through programmability. In fact,
Table~\ref{table:objclasses} categorizes existing interfaces and we clearly see
a trend towards reusable services.

@@

The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

Our approach is to expose more of the commonly used, code-hardened subsystems
of the underlying storage system as interfaces. The intent is that these
interfaces, which can be as simple as a redirection to the persistent data
store or as complicated as a strongly consistent directory service, should be
used and re-used in many contexts to implement a wide range of services. By
making programmability a `feature', rather than a `hack' or `workaround', we
help standardize a development process that is largely ad-hoc.

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-seq-thruput.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
load balancing interface alleviates load for multiple sequencers in
resource-constrained environments.  ``Underloaded (Baseline)'' is the
throughput of one sequencer running alone on a cluster; the other curves show
how load balancing affects the throughput of one sequencer when sequencers for
other logs are initially co-located on the same physical server (i.e. an
"Overloaded" environment).}\label{fig:mantle-seq-thruput}
\end{figure}

@@

We use a balancer policy specifically designed for multiple sequencers. The
balancer aggressively sheds half its load to the next server as soon as it can.
It is based off the ``Greedy Spill Even" approach in the
Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
Figure~\ref{listing:greedy-spill} for the reader's convenience. Each line of
code uses variables exposed by the Mantle API and is executed as a callback by
Mantle.  ``Server load" lets the administrator control what metric constitutes
load on a server; ``i" is a variable used by Mantle to iterate over servers and
``req\_rate" is a metric based on the exponenential decay of the request rate.
``When policy" is used by Mantle to set a conditional for kicking off migration;
in this case, the cluster migrates load if the current server, ``whoami", has
load (\texttt{>0.1}) and the next server (physical servers have ``ranks") has a
small amount load (\texttt{<0.1}). Finally, the ``where policy" is used by
Mantle to determine how much load to shed. The policy fills in a ``targets"
array to tell the migrator how much load to send to each server; in this case,
we shed half of the current load to the next server. The unit of load is a
single sequencer for a ZLog log.

\begin{figure}[]
\begin{lstlisting}
  -- Server load
  load = server[i]["req_rate"]

  -- When policy
  if server[whoami]["load"]>.01 and 
    server[whoami+1]["load"]<.01 then

  -- Where policy
  targets[whoami+1]=load/2
\end{lstlisting}
\caption{The sequencer balancing policy that works best for bursty workloads. 
\label{listing:greedy-spill}}
\end{figure}

@@

Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
sequencers as the load balancing interface migrates sequencers around the
cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
time (\(x\) axis).  ``Underloaded (Baseline)'' is the performance of one sequencer
on one physical server serving 7 clients - there are no other sequencers or
logs in the system and the curve represents the maximum throughput of a
sequencer on an underutilized physical server.  ``Overloaded: No Balancing" adds a
second sequencer of equal load to the same physical server with the intent of
overloading the single physical server; recall that the throughput for this
second sequencer is omitted. Compared to the baseline, the overloaded physical
server with 2 sequencers has more than a 2\(\times\) throughput degradation.
The stability of throughput curves (i.e. the variation on the \(y\) axis) indicate which sequencers
are on overloaded physical servers.  ``Underloaded (Baseline)" is the best
performance for this particular workload because there is no contention.
Performance is stable and the MDS is under-loaded because we do not see the
spike at the end like the other curves.  ``Overloaded: No Balancing" shows an
overloaded sequencer because the performance of the monitored sequencer is low
and unstable (high variation in performance).

The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
improvements when sequencers are load balanced across the cluster.
The load balancing policies differ by the threshold at which they start balancing:
``Overloaded: Sequencer Policy" starts balancing immediately and ``Overloaded: Conservative
Policy" waits for 120 seconds before migrating sequencers.

>>>>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
its code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to make trade-offs for latency/throughput using lease capabilities, balance load, propogate
interfaces, and recover from failure.
\end{abstract}

@@

A number of recent developments are disrupting traditional storage systems.
First, the falling prices of flash storage and the availability of new types of
non-volatile memory that are orders of magnitude faster than traditional
spinning media are moving overall performance bottlenecks away from storage
devices to CPUs and networking, and pressure storage systems to shorten their
code paths and incorporate new
optimizations~\cite{gray_tape_2007,gray_flash_2008}. 
Second, the demand for
managing structured data and flexible consistency semantics at scale pressure
big data processing systems to use storage abstractions that can meet these
demands~\cite{apache_contributors_parquet_2014}. 
Finally, production-quality
scalable storage systems available as open source software have established and
are continuing to establish new, \emph{de-facto} API standards at a faster pace
than traditional standards
bodies~\cite{snia_implementing_2014,linux_foundation_kinetic_2015}.

@@

These three trends put evolutionary pressure on storage systems and raise the
question of whether there are principles that storage systems designers can follow
to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.
% that are important for
%users to continue to entrust their data to the storage system.

In this paper we investigate an approach that focuses on generalizing and
composing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstractions, one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency. 
%We will refer to this programmability as
%\emph{programmable storage}, which differs from \emph{active storage} (the
%injection and execution of arbitrary code in a storage system) 
%and
%\emph{software-defined storage} (the control of thin-provisioning of storage).

{\it \textbf{We define a programmable storage system}} to be a storage system that facilitates the 
re-use and extension of existing storage abstractions and services provided by the underlying software stack,
to enable the creation of new storage abstractions via composition.
A programmable storage system is realized by exposing existing functionality (such as 
metadata services and synchronization capabilities) as interfaces that can be ``glued together''
using a high-level language.  Programmable storage differs from \emph{active storage},
or the injection and execution of arbitrary code in a storage system.

To illustrate the benefits and challenges of this approach we have designed and
evaluated Malacology, a programmable storage system for constructing new
services by re-purposing existing subsystem abstractions of the storage stack.
We build the framework in Ceph, a popular, open source software storage stack.
We choose Ceph to demonstrate the concept of programmable stoarge because it
offers a broad spectrum of existing services, including distributed locking and
caching services provided by metadata servers, durability and logical storage
devices provided by the backend object store, and propagation of consistent
cluster state provided by the monitoring service (see
Figure~\ref{fig:overview}). As we will show in this paper, this framework is
expressive enough to provide the functionality necessary for implementing new
services. Our contributions are:

Our prototype includes a non-exhaustive set of interfaces that can be used as building blocks
for constructing novel storage abstractions, including:

@@

\item Versioning interface that uses the underyling storage stack's consensus
infrastructure

\item Logical Device interface for adding functionality to daemons that manage
storage/metadata devices, allowing users to push computation to the data

\item An interface for mediating access to a shared resources, which exposes
serialization primatives using capabilities, caching, and batching

\item Load Balancing interface for managing resources across the cluster in
multi-tenant environments

\item Durability interface for persisting policies using the underyling storage
stack's object store

@@

\item A high-performance distributed shared log service called ZLog which is
based on CORFU~\cite{balakrishnan_corfu_2012}

@@

\section{Highly Tailored, Application-Specific Storage Stacks
}\label{highly-tailored-and-application-specifc-storage-systems}

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by exposing
tunable parameters to control low level implementation details like block size.
Unfortunately, users want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

@@

To demonstrate this trend in storage systems, we examine Ceph's
programmability. Something of a storage swiss army knife, Ceph supports file,
block, and object interfaces simultaneously in a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
programmability; the OSDs support the implementation of domain-specific logical
object interfaces by composing existing interfaces. These interfaces are
written in C++ and are statically loaded into the system.

@@

The Ceph community provides empirical evidence that developers embrace
programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
growth in the production use of domain-specific interfaces in the Ceph
community since 2010.  What is most remarkable is that this trend contradicts
the notion that API changes are a burden for users.  Rather it appears that a
gap in existing interfaces are being addressed through programmability. In
fact, Table~\ref{table:objclasses} categorizes existing interfaces and we
clearly see a trend towards reusable services.

@@

The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to achieve programmability because their needs, whether they be
related to performance, availability, consistency, correctness etc., are not
satisfied by the underlying storage system. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custome object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

Our approach is to expose more of the commonly used, code-hardened subsystems
of the underlying storage system as interfaces. The intent is that these
interfaces, which can be as simple as a redirection to the persistent data
store or as complicated as a strongly consistent directly service, should be
used and re-used in many contexts to implement a wide range of services. By
making programmability a `feature', rather than a `hack' or `workaround', we
help standardize a development process that is largely ad-hoc.

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-seq-thruput.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}]
Performance profiles of a single sequencer running in a resource constrained
environment show the benefits of using the load balancing interface.  ``1
Sequencer Baseline'' is the throughput of one sequencer on an underloaded
physical server; the other curves show how load balancing affects the
throughput of one sequencer when sequencers for other logs are co-located on
the same physical server.}\label{fig:mantle-seq-thruput}
\end{figure}

@@

We use a balancer policy specifically designed for multiple sequencers. The
balancer aggressively sheds half its load to the next server as soon as it can.
It is based off the ``Greedy Spill Even" approach in the
Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
Figure~\ref{listing:greedy-spill} for the reader's convenience. An MDS is a
physical server capable of housing a sequencer.

\begin{figure}[]
\begin{lstlisting}
  -- Metadata server load
  mdsload = MDSs[i]["all"]

  -- When policy
  if MDSs[whoami]["load"]>.01 and 
    MDSs[whoami+1]["load"]<.01 then

  -- Where policy
  targets[whoami+1]=allmetaload/2
\end{lstlisting}
\caption{The sequencer balancing policy that works best for bursty workloads.
\label{listing:greedy-spill}}
\end{figure}

@@

Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
sequencers as the load balancing interface migrates sequencers around the
cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
time (\(x\) axis).  ``1 Sequencer Baseline'' is the performance of one sequencer
on one physical server serving 7 clients - there are no other sequencers or
logs in the system and the curve represents the maximum throughput of a
sequencer on an underutilized physical server.  ``No Balancing" adds a
second sequencer of equal load to the same physical server with the intent of
overloading the single physical server; recall that the throughput for this
second sequencer is omitted. Compared to the baseline, the overloaded physical
server with 2 sequencers has more than a 2\(\times\) throughput degradation.
The stability of throughput curves (i.e. the variation on the \(y\) axis) indicate which sequencers
are on overloaded physical servers.  ``1 Sequencer Baseline" is the best
performance for this particular workload because there is no contention.
Performance is stable and the MDS is under-loaded because we do not see the
spike at the end like the other curves.  ``No Balancing" shows an
overloaded sequencer because the performance of the monitored sequencer is low
and unstable (high variation in performance).

The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
improvements when sequencers are load balanced across the cluster. Both curves
show the throughput of one of the sequencers when a second sequencer is
running on the same physical server. The load balancing policies differ by the threshold at which they start balancing:
``Sequencer Policy" starts balancing immediately and ``Conservative
Policy" waits for 120 seconds before migrating sequencers.
